---
title: 'Regression Case Study: Boston Housing'
author: "Michael Harrison"
date: "May 1, 2017"
output: html_document
---


1. CRIM: per capita crime rate by town
2. ZN: proportion of residential land zoned for lots over 25,000 sq.ft.
3. INDUS: proportion of non-retail business acres per town
4. CHAS: Charles River dummy variable (= 1 if tract bounds river; 0 otherwise) 
5. NOX: nitric oxides concentration (parts per 10 million)
6. RM: average number of rooms per dwelling
7. AGE: proportion of owner-occupied units built prior to 1940
8. DIS: weighted distances to five Boston employment centers
9. RAD: index of accessibility to radial highways
10. TAX: full-value property-tax rate per $10,000
11. PTRATIO: pupil-teacher ratio by town
12. B: 1000(Bk − 0.63)2 where Bk is the proportion of blacks by town 
13. LSTAT: % lower status of the population
14. MEDV: Median value of owner-occupied homes in $1000s

# 1. Prepare Problem
# a) Load packages
```{r}
library(mlbench)
library(caret)
library(corrplot)
```

# b) Load dataset
```{r}
data("BostonHousing")
```

# c) Split-out validation dataset
```{r}
inTrain <- createDataPartition(BostonHousing$medv, p=0.8, list=FALSE)
training <- BostonHousing[inTrain,]
testing <- BostonHousing[-inTrain,]
```

# 2. Summarize Data
# a) Descriptive statistics
```{r}
dim(training)
```
```{r}
sapply(training, class)
```

```{r}
head(training, 6)
```
```{r}
summary(training)
```

```{r}
training[,4] <- as.numeric(as.character(training[,4]))
cor(training[,1:13])
```

# b) Data visualizations

```{r}
par(mfrow = c(3,5))
for(i in 1:13){
        hist(training[,i], main = names(training)[i], xlab = names(training)[i])
}

```

# 3. Prepare Data
# a) Data Cleaning
# b) Feature Selection
# c) Data Transforms

# 4. Evaluate Algorithms
# a) Test options and evaluation metric
# b) Spot-Check Algorithms
# c) Compare Algorithms

# 5. Improve Accuracy
# a) Algorithm Tuning
# b) Ensembles

# 6. Finalize Model
# a) Predictions on validation dataset
# b) Create standalone model on entire training dataset
# c) Save model for later use